ohun: an R package for diagnosing and optimizing automatic acoustic signal detection

### 2022-07-27 manuscript/manuscript.Rmd
update ohun paper

---
title: 'ohun: an R package for optimizing automatic acoustic signal detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: combined_bibs.bib
citation-style: apa.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc")

@@

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."
opts_knit$set(root.dir = rootdir)

@@

```{r, fixing citation .bib files, eval = TRUE, echo = FALSE}

# how to cite in Rmarkdown
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

# Here include the path to the libraries for each author. This code will copy/update the library each time the Rmarkdown file is compiled
# Alternatively authors ca
personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib")

# update bibtex library
for (i in 1:length(personal_libraries))
if (file.exists(personal_libraries[i]))
 file.copy(from = personal_libraries[i], to = file.path(dirname(whereami::thisfile()), paste0(names(personal_libraries)[i], ".bib")), overwrite = TRUE)

## combine .bib files in compiled.bib ###
# list bib files
bibs <- list.files(pattern = ".bib$", path = dirname(whereami::thisfile()), full.names = TRUE)

# remove combined from pool
bibs <- bibs[basename(bibs) != "combined_bibs.bib"]

#  read in a list
if (length(bibs) > 0){
  combined_bibs_l <- lapply(bibs, readLines)
  names(combined_bibs_l) <- gsub("\\.bib$", "", basename(bibs))

# combine bibs in a single one
combined_bibs <- unlist(combined_bibs_l)

# save pool libraries as combined_bibs.bib
writeLines(text = combined_bibs, file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

# check if there are duplicated keys
keys <- grep("@article{", combined_bibs, fixed = TRUE, value = TRUE)
keys <- gsub("@article{", "", keys, fixed = TRUE)
keys <- gsub(",", "", keys, fixed = TRUE)

tab_keys <- table(keys)

if (anyDuplicated(keys)){
  print(paste0(sum(tab_keys > 1), " duplicate(s) references found in combined_bibs.bib"))
}
} else writeLines(text = "empty", file.path(dirname(whereami::thisfile()), "combined_bibs.bib"))

@@

Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
Grace Smith-Vidaurre ^4^,
Gloriana Chaverri ^3^,
Fabiola Chirino ^1^ &
Alejandro Rico-Guevara ^5^

@@

^5^  Department of Biology, University of Washington, Seattle, USA

@@

### Keywords:

@@

Animal acoustic signals are widely used in a diversity of research areas. This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings. However, bioacoustic research can easily generate large amounts of data which might proof challenging to analyze in a timely manner. Many tools for the automatic detection of signals are currently available. However, choosing the right approach for a specific task might turn difficult as there are no available tools that can provide a common framework for evaluating detection performance. Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines as well as compare the performance of different detection approaches. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices. This can be done with routine outputs imported from other software and from detection run within the package itself. The package also provides a set functions to organize acoustic data sets in a format amenable for detection analyses. It also includes an implementation of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We included examples on how to evalute detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls. Finally, we provide some general suggestions to improve detection performance.

@@

Animal acoustic signals are widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive. In addition, the existence of online repositories and growing number has facilitated the study of these traits at larger taxonomic and geographic scales. However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e. lots of recordings), which can be challenge to analyzed manually by human observers. As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewed by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.

@@

The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of tools also posits a challenge as it can be hard to navigate. In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system. The performance of automatic acoustic signal detection routines have been evaluated using standard indices from signal detection theory (CITATIONS). In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated. This comparison allows to quantify the number of correctly detected signals (true positives), wrongly detected signals (false positives) and missed signals (false negatives) as well as additional metrics derived from these indices (e.g. recall, precision). However, the fact that acoustic signals are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures) demand additional information to fully diagnose detection performance. This is particularly relevant when the precise time position of signals is needed as is common for research in which measures of the acoustic structure of signals is the main goal. For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one. Therefore, metrics that account for this other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.

Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose specific aspects of acoustic detection routines in order to simplify their optimization. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines. This can be done with routine outputs imported from other software as well as detection run within the package itself. The package also provides a set functions to explore acoustic data sets and organize them in a format amenable for detection analyses. In addition, it offers implementations of two automatic detection methods commonly used in bioacustic analysis: energy-based detection and template-based detection. We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluate and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).

@@

In this case all recordings have format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel). We can also check the duration and size of files. This information is important as some tuning parameters of detection routines can behave differently depending on file format (e.g. time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats. In addition, long sound files might be hard to analyzed in regular computers and might have to be split in shorter clips. In the latter case the function `split_acoustic_data` can be used to produce those clips:

@@

A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1. However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted. The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study. These indices provide an useful framework for diagnosing and optimizing the performance of a detection routine.

The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above. To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files. For instance, the following code evaluates a routine run in Raven Pro 1.6 (XXXX) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.54989 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:

@@

raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = "Raven.Band.Limited.Energy.Detector.selections.txt", warbler.format = TRUE)

# raven_ref <- manual_ref[manual_ref$sound.files %in% raven_detec$sound.files, ]

@@

diagnose_detection(reference = manual_ref_tae, detection = raven_detec)

@@

The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 6 files:

@@

diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by.sound.file = TRUE)

head(diag_raven)
```

Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection. The process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.

@@

The package offers two methods for automatic signal detection: template-based and energy-based detection. These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively. If the target signals do not fit these requirements, more elaborated methods (i.e. machine/deep learning approaches) are warranted.

@@

We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats at Baru Biological Station, in south-western Costa Rica in January 2020. Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants). Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software. Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6 (XXXX). Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals. A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording). Annotations were made with a time window of 200 samples and 70% of overlap. Annotations were then imported into R using the package Rraven (XXXX).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped. Most variation is found between individuals although the basic form of a short, downward broadband frequency modulation is always shared (Fig. BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

@@

# knitr::include_graphics("~/Dropbox/Projects/ohun_paper/data/raw/thyroptera/Catalog_p1.jpeg")
knitr::include_graphics((file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg")))
```

Template-based detection is a useful approach when there is little structural differences in the target signals. We used this approach in ohun to detect inquiry calls. To do this, we tested the performance of three acoustic templates on a training subset of five sound files. The function `get_templates` finds several signals representative of the variation in signal structure. This function measures several spectrographic parameters which are then summarized using Principal Component Analysis. The first two components are used to project the acoustic space. On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space. Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space. The user needs to define the number of sub-spaces in which the acoustic space will be split.

@@

```{r, echo = FALSE, fig.height= 6, fig.cap= "FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first two component of a Principal Component Analysis on spectrographic parameters. Templates are selected as those closer to the centroid for each of the sub-spaces. Gray dashed lines delimit the region of sub-spaces. Yellow circles around points highlight the position of the signals selected as templates."}

@@

The last line of codes evalutes the detection on the test data set, which shows a good performance for both recall and precision.

### Energy-based detection on zebra finch songs

We used recordings from 18 zebra finch males. These recordings were obtained ....

Zebra finch vocalizations are composed by multiple elements (i.e. distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO). However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.  
```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. ZEBRAFINCH-SPECTRO. Example spectrograms of male zebra finch songs for each of the 18 sound files used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 359 ms and the frequency range 0-11 kHz. Signals have been highlighted for visualization purposes only.", error=TRUE}

knitr::include_graphics((file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg")))

@@

The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
```{r, eval = TRUE, echo = FALSE}

@@

The output (`opt_det_train`) shows the performance indices for each of those combinations, here we show the 10 combinations with the highest F1 score:

@@

The performance on the test data set was also acceptable with a F1 score of 0.95. Note that in the example we use a small subset of sound files for training. More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.

@@

The package offers additional tools to simplify acoustic signal detection. The function 'feature_reference' allows to extract information about the duration and frequency range of acoustic signals than can then be used to define tuning parameter values. This can be particularly useful for energy-based detection in which time and frequency attributes of the target signals are required. Detected signals can be labeled as false or true positives with the function 'label_detection'. This allows users to explore the structure of false positives and figure out ways to exclude them. The function 'filter_detections' can be used to remove ambiguous signals (i.e. those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e. the highest template correlacion). Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).

@@

Here we have showed how to evaluate the performance of acoustic signal detection routines using the package ohun. The package can evaluate detection outputs imported from other software, as well as its own detection routines. The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection. Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform. Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question. Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data. Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation. For instance detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram).

The package provides two native detection methods: template-based  and energy-based detection. Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools. However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases. Deep learning methods tend to require larger computational power and more complex training routines. This might bring unnecessary difficulties when dealing when less challenging detection tasks. Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals. Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algortihms. Acoustic parameters can be used to quantify the structure of the signals and tell apart target from non-target signals.

Detection routines can take a long time when working with large amounts of acoustic data (e.g. large recordings and/or many files). These are some tips that can help make a routine more time-efficient. 1. Always test procedures on small data subsets. Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis. 2. Template-based detection is almost always faster than energy-based detection. 4. Run routines in parallel. Parallelization (i.e. the ability to distribute tasks over several cores in your computer) can significantly speed-up routines. All functions for automatic detection and performance evaluation allow users to run analysis in parallel (see `parallel` argument in those functions). Hence a computer with several cores can be helpful for improving efficiency. 4. Sampling rate matters. Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals. 5. Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data. These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.

>>>>

---
title: 'ohun: an R package for diagnosing and optimizing automatic acoustic signal detection'
author:
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float:
      collapsed: no
      smooth_scroll: no
  word_document:
    toc: yes
    toc_depth: '3'
  pdf_document:
    toc: yes
    toc_depth: '3'
always_allow_html: yes
bibliography: references.bib
citation-style: methods-in-ecology-and-evolution.csl
link-citations: yes
editor_options:
  chunk_output_type: console
---

@@

pkgs <- c("kableExtra", "knitr", "rprojroot", "sciRmdTheme", "rmdwc", "whereami", "maRce10/ohun", "ggplot2", "viridis", "rmdwc", "RefManageR", "bibtex")

@@

# set working directory as project directory or one directory above,
rootdir <- try(rprojroot::find_rstudio_root_file(), silent = TRUE)
if (is(rootdir, "try-error")) rootdir <-  ".."

opts_knit$set(root.dir = rootdir)

@@

```{r, fixing citation .bib files, eval = TRUE, echo = FALSE, message=FALSE, warning=FALSE}

# how to cite in Rmarkdown:
# https://rmarkdown.rstudio.com/authoring_bibliographies_and_citations.html

# find citation style files (csl) here:
# https://github.com/stackpr/csl-styles

# Here include the path to the libraries for each author. This code will copy/update the library each time the Rmarkdown file is compiled
# Alternatively authors ca
personal_libraries <- c(author_1 = "/home/m/Documents/mendeley_libraries/ohun_paper.bib", author_2 = "/home/m/Documentos/mendeley_bibtex_files/ohun_paper.bib")

# location of current Rmarkdown
rmd_dir <- dirname(whereami::thisfile()) 
if (is.null(rmd_dir)) rmd_dir <- file.path(rootdir, "manuscript")

# read personal library
personal_libraries <- personal_libraries[file.exists(personal_libraries)]
personal_bib <- suppressMessages(ReadBib(personal_libraries))

# read paper's references.bib if it exists
if (file.exists(file.path(rootdir, "references.bib"))){
    combined_bib <- suppressMessages(ReadBib(file.path(rmd_dir, "references.bib")))

    merge_bib <- suppressMessages(merge(personal_bib, combined_bib, fields.to.check = "title", ignore.case = TRUE))
} else
    merge_bib <- personal_bib

# save combined references
WriteBib(bib = merge_bib, file = file.path(rmd_dir, "references.bib"), biblatex = TRUE)

@@

Marcelo Araya-Salas ^1,^ ^2,^ ^3^ *,
Grace Smith-Vidaurre ^4^,
Gloriana Chaverri ^3^, ^5^ 
Fabiola Chirino ^1^ &
Alejandro Rico-Guevara ^6^

@@

^5^  Smithsonian Tropical Research Institute, Panama City, Panamá

^6^  Department of Biology, University of Washington, Seattle, USA

@@

**Keywords**: bioacoustics, automatic detection, animal vocalizations

@@

Animal acoustic signals are widely used in a diversity of research areas. This is partly due to the relative ease in which they can be registered for wide range of taxonomic groups and research settings. However, bioacoustic research can easily generate large amounts of data which might prove challenging to analyze in a timely manner. Many tools for the automatic detection of signals are currently available. However, choosing the right approach for a specific task might turn difficult as there are no available tools that can provide a common framework for evaluating detection performance. Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, by providing functions to diagnose and optimize detection routines as well as compare the performance of different detection approaches. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines, using common signal detection theory indices. This can be done with routine outputs imported from other software and from detection run within the package itself. The package also provides a set of functions to organize acoustic data sets in a format amenable for detection analyses. It also includes an implementation of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection. We included examples on how to evaluate detections from external software and showcase the package usage with study cases on zebra-finch (*Taenopygia gutata*) songs and Spix's disc-winged bat (*Thyroptera tricolor*) ultrasonic social calls. Finally, we provide some general suggestions to improve detection performance.

@@

Animal acoustic signals are widely used to investigate a variety of questions in highly diverse areas, ranging from neurobiology to community ecology and evolutionary biology. The wide usage of animal sounds in research is partly due to the fact that they can be easily registered using non-intrusive methods, can be obtained in a variety of settings from laboratories to natural areas, and the equipment required for registering an analyzing these signals has become increasingly inexpensive. In addition, the existence of online repositories has facilitated the study of these traits at larger taxonomic and geographic scales. However, the adoption of bioacustic approaches in research can also imply large amounts of data (i.e. lots of recordings), which can be challenging to analyzed manually. As a result, a growing number of computational tools for the analysis of acoustic features on those signals is increasingly available (reviewed by @Stowell2022), reflecting the need of automated approaches for efficiently conducting these analyses.

@@

The growing availability of tools for automatic detection of acoustic events, particularly as free software, is expected to further simplify acoustic data processing, making it accessible to a wider range of users and scientific questions. However, this diversity of tools also posits a challenge as it can be hard to navigate. In this regard, the use of standard approaches for evaluating the performance of automatic detection tools might prove helpful to inform the decision about which method better fit our study system. The performance of automatic acoustic signal detection routines has been evaluated using standard indices from signal detection theory (CITATIONS). In its basic form, performance is assessed by comparing the output of a detection routine against a "gold standard" reference in which all the target signals have been annotated. This comparison allows to quantify the number of correctly detected signals (true positives), wrongly detected signals (false positives) and missed signals (false negatives) as well as additional metrics derived from these indices (e.g. recall, precision). However, the fact that acoustic signals are not being evaluated as discrete classification units (as opposed to for instance identifying species from pictures) demand additional information to fully diagnose detection performance. This is particularly relevant when the precise time position of signals is needed as is common for research in which measures of the acoustic structure of signals is the main goal. For instance, the same signal can be detected as several separated signals, the inferred time position can be far off from the the target signal position, or several signals can be detected as a single one. Therefore, metrics that account for these other performance dimensions can be a valuable tool to properly diagnose automatic acoustic signal detection.

Here we present the new R package ohun. The package is intended to facilitate the automatic detection of acoustic signals, providing functions to diagnose specific aspects of acoustic detection routines in order to simplify their optimization. The package makes use of reference annotations containing the time position of target signals in a training data set to evaluate the performance of detection routines. This can be done with routine outputs imported from other software as well as detection run within the package itself. The package also provides a set of functions to explore acoustic data sets and organize them in a format amenable for detection analyses. In addition, it offers implementations of two automatic detection methods commonly used in bioacoustic analysis: energy-based detection and template-based detection. We explain how to explore and format acoustic data sets, how acoustic signal detection routines can be evaluated and then showcase the package usage with study cases on zebra-finch songs (*Taenopygia gutata*) and Spix's disc-winged bats (*Thyroptera tricolor*) which represent different recording settings (lab and flight cages) and signal types (sonic mating signals and ultrasonic social calls).

@@

In this case all recordings have format (.wav files, 44.1 kHz sampling rate, 16 bit resolution and a single channel). We can also check the duration and size of files. This information is important as some tuning parameters of detection routines can behave differently depending on file format (e.g. time window size is affected by sampling rate) or simply because some software might can only work on specific sound file formats. In addition, long sound files might be hard to analyze in regular computers and might have to be split in shorter clips. In the latter case the function `split_acoustic_data` can be used to produce those clips:

@@

A perfect detection will have no false positives or false negatives, which will result in both recall and precision equal to 1. However, perfect detection cannot always be reached and some compromise between detecting all target signals plus some noise (recall = 1 & precision < 1) and detecting only target signals but not all of them (recall < 1 & precision = 1) might be warranted. The right balance between these two extremes will be given by the relative costs of missing signals and mistaking noise for signals, given the specific goals of the study. These indices provide a useful framework for diagnosing and optimizing the performance of a detection routine.

The package offers tools to evaluate the performance of an acoustic signal detection based on the indices described above. To accomplish this, annotations derived from a detection routine are compared against a reference table containing the time position of all target signals in the sound files. For instance, the following code evaluates a routine run in Raven Pro 1.6 (Cornell Lab of Ornithology) using the "band limited energy detector" option (minimum frequency: 0.8 kHz; maximum frequency: 22 kHz; minimum duration: 0.03968 s; maximum duration: 0.54989s; minimum separation: 0.02268 s) on a subset of the Zebra finch recordings described below:

```{r read and save raven automatic detection selection files, eval = FALSE, echo = FALSE}

@@

raven_detec <- imp_raven(path = file.path(rootdir, "/data/raw/"), files = c("22.kHz_sep.0.02322.selections.txt.txt", "15.kHz_sep.0.02322.selections.txt", "10.kHz_sep.0.02322.selections.txt", "10.kHz_sep.0.005.selections.txt"), warbler.format = TRUE)

raven_detec$tuning_parameters <- gsub(".selections.txt", "", raven_detec$selec.file)
raven_detec$tuning_parameters <- gsub("kHz_", "kHz; ", raven_detec$tuning_parameters)
raven_detec$tuning_parameters <- gsub(".kHz", " kHz", raven_detec$tuning_parameters)
raven_detec$tuning_parameters <- gsub("sep.", "sep: ", raven_detec$tuning_parameters)
raven_detec$tuning_parameters <- paste0("band: 1-", raven_detec$tuning_parameters, " s")

write.csv(raven_detec, "./data/processed/combined_raven_detection.csv", row.names = FALSE)

raven_detec <- read.csv("combined_raven_detection.csv")

diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

@@

raven_detec <- read.csv("./data/processed/combined_raven_detection.csv")

diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters")

@@

The output shows the indices described above. The function also allows to detail those indices by sound file. Here we show the first 10 files:

@@

diag_raven <- diagnose_detection(reference = manual_ref_tae, detection = raven_detec, by = "tuning_parameters", by.sound.file = TRUE)

head(diag_raven, 10)
```

Diagnostics from routines using different tuning parameters can be used to identify the parameter values that optimize a detection. These process of evaluating different routines for detection optimization is incorporated into the two signal detection approaches provided natively by ohun, which we depict in the following section. Note that the detection with Raven Pro does not necessarily reflect the best performance of this software and has been included only as an example on evaluating detection from external sources rather than a direct comparison of performance to ohun.

@@

The package offers two methods for automatic signal detection: template-based and energy-based detection. These methods are better suited for highly stereotyped or good signal-to-noise ratio (SNR) signals, respectively. If the target signals do not fit these requirements, more elaborate methods (i.e. machine/deep learning approaches) are warranted.

@@

We recorded `r length(unique(manual_ref_thy$sound.files))` individuals of Spix's disc-winged bats (*Thyroptera tricolor*) at Baru Biological Station, in south-western Costa Rica in January 2020. Bats were captured at their roosting sites (furled leaves of Zingeberaceae plants). Each individual bat was released in a large flight cage (9 x 4 x 3 m) for a 5 min period and their ultrasonic inquiry calls were recorded using a condenser microphone (CM16, Avisoft Bioacoustics, Glienike/Nordbahn, Germany) through an Avisoft UltraSoundGate 116Hm plugged into a laptop computer running Avisoft-Recorder software. Recordings were made at a sampling rate of 500 kHz and an amplitude resolution of 16 bits.

Recordings were manually annotated using Raven Pro 1.6 (Cornell Lab of Ornithology). Annotations were created by visual inspection of spectrograms, in which the start and end of signals were determined by the location of the continuous traces of power spectral entropy of the target signals. A total of `r nrow(manual_ref_thy)` calls were annotated (~`r round(nrow(manual_ref_thy) / 30, 0)` calls per recording). Annotations were made with a time window of 200 samples and 70% of overlap, and were then imported into R using the package Rraven (@Araya-Salas2017e).

Inquiry calls of Spix's disc-winged bats are structurally stereotyped (@Chaverri2010b). Most variation is found between individuals, although the basic form of a short, downward broadband frequency modulation is always shared (Fig. BAT-SPECTRO, Araya-Salas et al 2021 ontogeny).

@@

knitr::include_graphics(file.path(rootdir, "data/raw/thyroptera/Catalog_p1.jpeg"))

Template-based detection is a useful approach when there are few structural differences in the target signals. We used this approach in ohun to detect inquiry calls. To do this, we tested the performance of three acoustic templates on a training subset of five sound files. The function `get_templates` finds several signals representative of the variation in signal structure. This function measures several spectrographic parameters which are then summarized using Principal Component Analysis. The first two components are used to project the acoustic space. On this space the function defines sub-spaces as equal-size slices of a circle centered at the centroid of the acoustic space. Templates are then selected as those closer to the centroid within each of the sub-spaces, including the centroid for the entire acoustic space. The user needs to define the number of sub-spaces in which the acoustic space will be split.

@@

```{r, echo = FALSE, fig.height= 6, fig.cap= "FIG. ACOUSTIC-SPACE. Acoustic space defined as the the first two components of a Principal Component Analysis on spectrographic parameters. Templates are selected as those closer to the centroid for each of the sub-spaces. Gray dashed lines delimit the region of sub-spaces. Yellow circles around points highlight the position of the signals selected as templates."}

@@

The last line of codes evaluates the detection on the test data set, which shows a good performance for both recall and precision.

### Energy-based detection on zebra finch vocalizations

We used recordings from 18 zebra finch males recorded at obtained from the Rockefeller University Field Research Center Song Library (http://ofer.sci.ccny.cuny.edu/songs, @Tchernichovski2021). Recordings contain undirected vocalizations (songs or calls) of single males recorded in sound attenuation chambers using Sound Analysis Pro. Zebra finch vocalizations are composed by multiple elements (i.e. distinct patterns of continuous traces of power spectral entropy in the spectrogram separated by time gaps) that can vary importantly in key features as duration and frequency range (ZEBRAFINCH-SPECTRO). However, as the recorded signals show a good signal-to-noise ratio, they can potentially be detected using an energy-based approach.  
```{r catalog zebra f, out.width = "100%", echo = FALSE, fig.align= "left",fig.cap="Fig. ZEBRAFINCH-SPECTRO. Example spectrograms of male zebra finch songs for each of the 18 sound files used in the analysis. The highest signal-to-noise ratio call by sound file are shown. The time scale range is 359 ms and the frequency range 0-11 kHz. Signals have been highlighted for visualization purposes only.", error=TRUE}

knitr::include_graphics(file.path(rootdir, "data/raw/taeniopygia/same.time-Catalog_p1.jpeg"))

@@

Reference annotations were made manually on the oscillogram with the spectrogram and audio as a guide using Raven Lite 2.0.1 (Cornell Lab of Ornithology). The following code loads the reference annotations and split them in two data set for training (3 sound files) and testing (15 sound files):
```{r, eval = TRUE, echo = FALSE}

@@

The output (`opt_det_train`) shows the performance indices for each of those combinations. Hhere we show the 10 combinations with the highest F1 score:

@@

The performance on the test data set was also acceptable with a F1 score of 0.95. Note that in the example we used a small subset of sound files for training. More training data might be needed for optimizing a detection routine on larger data sets or recordings with more variable signals or background noise levels.

@@

The package offers additional tools to simplify acoustic signal detection. Detected signals can be labeled as false or true positives with the function 'label_detection'. This allows users to explore the structure of false positives and figure out ways to exclude them. The function 'filter_detections' can be used to remove ambiguous signals (i.e. those labeled as split or merged detection) keeping only those that maximize a specific criterium (i.e. the highest template correlacion). Finally, note that  several templates representing the range of variation in signal structure can be used to detect semi-stereotyped signals when running template-based detection ('template_detection' function).

@@

Here we have shown how to evaluate the performance of acoustic signal detection routines using the package ohun. The package can evaluate detection outputs imported from other software, as well as its own detection routines. The latter can be iterated over combinations of tuning parameters in order to find those values that optimize a detection. Despite signal detection indices being commonly reported when presenting new automatic detection methods, to our knowledge widely applicable performance evaluating routines have not been made available in a free, open source platform. Providing a common framework for acoustic signal detection evaluation can simplify comparing performances of different tools and the selection of those better suited to our system and research question. Note that the tools offered by ohun for diagnosing a detection should not be necessarily limited to acoustic data. Any type of automatic detection in which the time of occurrence of discrete events needs to be found can be evaluated and optimized by comparing it to a reference annotation, for instance, detection of specific behaviors in video analysis of animal motor activity (Sturman et al 2020; Hsu & Yttri 2021; DeepEthogram). The package can also be used to assess the repeatability of manually made annotations by comparing annotations from different observers on the same data sets.

The package provides two native detection methods: template-based  and energy-based detection. Compared to new deep learning approaches for finding the occurrence of acoustic signals, the two native methods are relatively simple tools. However, these methods have been widely used by the bioacoustic community and, under the appropriate conditions can reach adequate performance, as we have shown in our two study cases. Deep learning methods tend to require larger computational power, larger training data sets and, in some cases, more complex training routines. This might bring unnecessary difficulties when dealing with when less challenging detection tasks. Therefore the availability of a wide range of approaches is desired in order to simplify finding the most appropriate tool for the intricacies of our study system and research goals. Note that the tools offered in ohun can be used in a pipeline in which detected signals are further classified using more elaborated discrimination algorithms. Acoustic parameters can be used to quantify the structure of the signals and distinguish target from non-target signals.

Detection routines can take a long time when working with large amounts of acoustic data (e.g. large recordings and/or many files). These are some tips that can help make a routine more time-efficient. 1. Always test procedures on small data subsets. Make sure you are getting decent results on a small subset of recordings before trying to scale up the analysis. 2. Template-based detection is almost always faster than energy-based detection. 3. Run routines in parallel. Parallelization (i.e. the ability to distribute tasks over several cores in your computer) can significantly speed-up routines. All functions for automatic detection and performance evaluation allow users to run analysis in parallel (see `parallel` argument in those functions). Hence a computer with several cores can be helpful for improving efficiency. 4. Sampling rate matters. Detecting signals on low sampling rate files is faster, so we must avoid having Nyquist frequencies much higher than the highest frequency of the target signals. 5. Try using a computer with lots of RAM memory or a computer cluster for working on large amounts of data. These rules are not restricted to ohun and can also be helpful to speed-up routines in other software packages.

There are some additional things to be considered when aiming to automatically detect acoustic signals. When running automatic signals detection try to use your knowledge about the signal structure to determine the initial range. This can be extremely helpful for narrowing down possible parameter values, particularly for energy-based detection. As a general rule, if people have a hard time figuring out where a target signal occurs, detection algorithms will also have a hard time. In cases in which occurrences are ambiguous low performances are expected. Make sure reference tables contain all target signals and only the target signals, otherwise performance optimization can be misleading as the performance of a detection cannot be better than the reference itself. Lastly, avoid having overlapping signals or several signals as a single one (like a multi-syllable vocalization) in the reference table when running an energy-based detector, as they are likely to be identified as separated units.

### Acknowledgements
